\bf Nonparametric Bootstrap KS Test with Stationary Time Series

### 2023-10-11 manuscript/manuscript.tex
removed underscores in text that prevented compiling
comments

@@

Consider a stationary time series $\{X_t: t = 1, \ldots, n\}$ with length~$n$. 
Our goal is to see if the data generating model of the series matches a 
hypothesized distribution. If the observations in the series were independent, a 
standard
nonparametric bootstrap KS test would draw a large number~$B$ bootstrap copies 
of the observed data, and compute the empirical distribution and a
fitted distribution from the fitted parameters
$\hat\theta_n^{(b)}$ for each copy $b = 1, \ldots, B$. ____ When serial 
dependence is present, the bootstrap procedure 
needs to preserve the serial dependence. Block bootstrap was motivated for this 
situation.

Consider the goodness of fit statistic:
\begin{equation*}
KS_n := \sup_x|Y_n(x; \hat\theta)|, 
Y_n(x; \hat\theta) = \sqrt{n}(F_n(x) - F(x; \hat\theta_n)),
\end{equation*}
where $F_n$ denotes the empirical distribution function based on $X_1,...,X_n$.
We note that
\begin{equation*}
Y_n(x; \hat\theta) = \sqrt{n}(F_n(x) - F(x)) - 
\sqrt{n}(F(x; \hat\theta_n) - F(x)),
\end{equation*}
where $F(x)$ is the true cdf (under the null $F(x) = F(x, \theta_0)$ for some
true parameter $\theta_0$).

Let us first consider the case where $X_i$'s are i.i.d. Denote by $F^*_n$ the
empirical distribution of the bootstrap sample $X^*_1,...,X^*_n$ and let
$\hat\theta^*_n$ be the parameter estimate based on the bootstrap samples. 
Using the bootstrap (asymptotic) theory, we can approximate the distribution of
\[
\sqrt{n}(F_n(x) - F(x))
\]
by that of
\[
\sqrt{n}(F^*_n(x) - F_n(x))
\]
and the distribution of
\[
\sqrt{n}(F(x; \hat\theta_n) - F(x))
\]
by that of
\[
\sqrt{n}(F(x; \hat\theta^*_n) - F(x; \hat\theta_n)).
\]
Therefore, if we define
\begin{equation*}
Y^*_n(x) = \sqrt{n}(F^*_n(x) - F_n(x)) - 
\sqrt{n}(F(x; \hat\theta^*_n) - F(x; \hat\theta_n)) 
= \sqrt{n}(F^*_n(x) - F(x; \hat\theta^*_n)) - 
\sqrt{n}(F_n(x) - F(x; \hat\theta_n)),
\end{equation*}
then $KS^*_n := \sup_x|Y^*_n(x)|$ is the bootstrap statistic that is expected
to approximate the distribution of $KS_n$. We note that the term
$\sqrt{n}(F_n(x) - F(x; \hat\theta_n))$ is exactly the bias term considered in 
Babu and Rao (2004).

We now consider the case where $X_i$'s are realizations from a time series and
$X^*_1,...,X^*_n$ are generated by block bootstrap. In this case, we can 
approximate the distribution of
\[
\sqrt{n}(F_n(x) - F(x))
\]
by that of
\[
\sqrt{n}(F^*_n(x) - E^*[F^*_n(x)])
\]
and the distribution of
\[
\sqrt{n}(F(x; \theta_n) - F(x))
\]
by that of
\[
\sqrt{n}(F(x; \hat\theta^*_n) - F(x; E^*[\hat\theta^*_n])).
\]
Here $E^*$ denotes the expectation with respect to the bootstrap distribution
(i.e., the randomness due to the resampling using block botostrap) conditional
on the observations $X_1,...,X_n$. We can compute $E^*[F^*_n(x)]$ and 
$E^*[\hat\theta^*_n]$ numerically (they can also be computed analytically, 
depending on the types of block bootstrap we use). For example, let 
$X^*_{1,b},...,X^*_{n,b}$ be the $b$th block bootstrap sample for 
$1 \leq b \leq B$. One can compute $F^*_{n,b}$ and $\hat\theta^*_{n,b}$ based on
$X^*_{1,b},...,X^*_{n,b}$. Then
\begin{equation*}
E^*[F^*_n(x)] \approx \frac{1}{B}\sum_{b = 1}^BF^*_{n, b}(x),
E^*[\hat\theta^*_n] \approx \frac{1}{B}\sum_{b = 1}^B\hat\theta^*_{n,b}.
\end{equation*}
In this case, we can define
\begin{equation*}
Y^*_n(x) = \sqrt{n}(F^*_n(x) - E^*[F^*_n(x)]) - 
\sqrt{n}(F(x; \hat\theta^*_n) - F(x; E^*[\hat\theta^*_n)]) 
= \sqrt{n}(F^*_n(x) - F(x; \hat\theta^*_n)) - 
\sqrt{n}(E^*[F^*_n(x)] - F(x; E^*[\hat\theta^*_n])),
\end{equation*}
and $KS^*_n = \sup_x|Y^*_n(x)|$.

\section{Null Hypothesis}
\label{sec:null}

\section{Alternative Hypothesis}
\label{sec:alternative}

>>>>

\jy{double blank liines to better separate paragraphs}

@@

Consider a stationary time series $\{X_t: t = 1, \ldots, n\}$ with length~$n$.
\jy{state the null hypothesis explicitly; we are testing the marginal
  distribution. Point out the challenges.}
Our goal is to see if the data generating model of the series matches a 
hypothesized distribution.

\jy{define $\hat\theta_n$}
If the observations in the series were independent, a 
standard
nonparametric bootstrap KS test would draw a large number~$B$ bootstrap copies 
of the observed data, and compute the empirical distribution and a
fitted distribution from the fitted parameters
$\hat\theta_n^{(b)}$ for each copy $b = 1, \ldots, B$. When serial 
dependence is present, the bootstrap procedure 
needs to preserve the serial dependence. Block bootstrap was motivated for this 
situation.

\jy{No double letter notations}
Consider the goodness of fit statistic:
\begin{equation*}
  T_n := \sup_x|Y_n(x; \hat\theta)|, 
Y_n(x; \hat\theta) = \sqrt{n}(F_n(x) - F(x; \hat\theta_n)),
\end{equation*}
where $F_n$ denotes the empirical distribution function based on $X_1,...,X_n$.
We note that
\begin{equation*}
Y_n(x; \hat\theta) = \sqrt{n}(F_n(x) - F(x)) - 
\sqrt{n}(F(x; \hat\theta_n) - F(x)),
\end{equation*}
where $F(x)$ is the true cdf (under the null $F(x) = F(x, \theta_0)$ for some
true parameter $\theta_0$).

Let us first consider the case where $X_i$'s are i.i.d. Denote by $F^*_n$ the
empirical distribution of the bootstrap sample $X^*_1,...,X^*_n$ and let
$\hat\theta^*_n$ be the parameter estimate based on the bootstrap samples. 
Using the bootstrap (asymptotic) theory, we can approximate the distribution of
\jy{no need for so many in-display equations}
\[
\sqrt{n}(F_n(x) - F(x))
\]
by that of
\[
\sqrt{n}(F^*_n(x) - F_n(x))
\]
and the distribution of
\[
\sqrt{n}(F(x; \hat\theta_n) - F(x))
\]
by that of
\[
\sqrt{n}(F(x; \hat\theta^*_n) - F(x; \hat\theta_n)).
\]
Therefore, if we define
\begin{equation*}
Y^*_n(x) = \sqrt{n}(F^*_n(x) - F_n(x)) - 
\sqrt{n}(F(x; \hat\theta^*_n) - F(x; \hat\theta_n)) 
= \sqrt{n}(F^*_n(x) - F(x; \hat\theta^*_n)) - 
\sqrt{n}(F_n(x) - F(x; \hat\theta_n)),
\end{equation*}
then $KS^*_n := \sup_x|Y^*_n(x)|$ is the bootstrap statistic that is expected
to approximate the distribution of $KS_n$. We note that the term
$\sqrt{n}(F_n(x) - F(x; \hat\theta_n))$ is exactly the bias term considered in 
Babu and Rao (2004).

We now consider the case where $X_i$'s are realizations from a time series and
$X^*_1,...,X^*_n$ are generated by block bootstrap. In this case, we can 
approximate the distribution of
\jy{No need for in-display}
\[
\sqrt{n}(F_n(x) - F(x))
\]
by that of
\[
\sqrt{n}(F^*_n(x) - E^*[F^*_n(x)])
\]
\jy{change the languages to add variety}
and the distribution of
\[
\sqrt{n}(F(x; \theta_n) - F(x))
\]
by that of
\[
\sqrt{n}(F(x; \hat\theta^*_n) - F(x; E^*[\hat\theta^*_n])).
\]
\jy{no closing a sentence before all the notations are introduced}
Here $E^*$ denotes the expectation with respect to the bootstrap distribution
(i.e., the randomness due to the resampling using block botostrap) conditional
on the observations $X_1,...,X_n$. We can compute $E^*[F^*_n(x)]$ and 
$E^*[\hat\theta^*_n]$ numerically (they can also be computed analytically, 
depending on the types of block bootstrap we use). For example, let 
$X^*_{1,b},...,X^*_{n,b}$ be the $b$th block bootstrap sample for 
$1 \leq b \leq B$. One can compute $F^*_{n,b}$ and $\hat\theta^*_{n,b}$ based on
$X^*_{1,b},...,X^*_{n,b}$. Then
\begin{align*}
E^*[F^*_n(x)] &\approx \frac{1}{B}\sum_{b = 1}^BF^*_{n, b}(x),\\
E^*[\hat\theta^*_n] &\approx \frac{1}{B}\sum_{b = 1}^B\hat\theta^*_{n,b}.
\end{align*}
In this case, we can define
\jy{too long}
\begin{equation*}
Y^*_n(x) = \sqrt{n}(F^*_n(x) - E^*[F^*_n(x)]) - 
\sqrt{n}(F(x; \hat\theta^*_n) - F(x; E^*[\hat\theta^*_n)]) 
= \sqrt{n}(F^*_n(x) - F(x; \hat\theta^*_n)) - 
\sqrt{n}(E^*[F^*_n(x)] - F(x; E^*[\hat\theta^*_n])),
\end{equation*}
and $KS^*_n = \sup_x|Y^*_n(x)|$.

\section{Simulation Studies}
\label{sec:simu}

\jy{First size and then power}

\section{Real Data Analysis}

\jy{Find some real time series to test for their marginal distributions. Open
  data, for example, temperatures or stock returns or ...}
